\section{Preparation}

\subsection{Lesson Plans}

This chapter covers the following ideas. When you create your lesson plan, it should contain examples which illustrate these key ideas. Before you take the quiz on this unit, meet with another student out of class and teach each other from the examples on your lesson plan. 

\input{10-SurfaceTripleIntegrals/SurfaceTripleIntegrals-Objectives}%$


%\subsection{Preparation Problems}

%Here are the preparation problems for this unit.

\subsection{Homework}

In the following list, the ``basic practice'' problems should be quick
problems to help you master the ideas.  The ``good problems'' will
require a little more work.  The theory and application problems are
ones that will challenge you more; make sure you do the problems from
this area to fully master the material.  

{\centering
\begin{tabular}{lcllll}\toprule
Topic &Sec &Basic Practice &Good Problems &Thy/App \\\midrule

Surface area & 14.5 & 1--24 & 27--34 & 37--41\\\hline
Parametric surfaces & 15.5 & 1--4, 35--42 & 13--26\\\hline
Surface integrals & 15.6 & 1--22 & 37--40\\\bottomrule
Triple integrals & 14.6 & 1--38, 63--66 & 43--58 & 39-40, 61--62, 67--70\\\hline
Spherical and Cylindrical & 14.7 & 1--22 & 23--38 & 43--46\\\hline
Jacobians & 14.8 & 27--30&&\\\hline
\end{tabular}

} 
\bigskip
It is crucial that you do not attempt to solve every integral.  For
the most part, you are learning to set up integrals in high
dimensions. You should do many more than 5 problems a day if you want
to become proficient at multiple integration.  \textbf{I would suggest
  that you do at least 15 or more problems a day in the assigned
  sections}, where you spend time setting up integrals and not solving
them.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../multivariable-calculus"
%%% End: 
